---
title: "Description and usage of MsBackenHmdb"
output:
    BiocStyle::html_document:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Description and usage of MsBackendHmdb}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    %\VignettePackage{Spectra}
    %\VignetteDepends{Spectra,BiocStyle}
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

**Package**: `r Biocpkg("MsBackendHmdb")`<br />
**Authors**: `r packageDescription("MsBackendHmdb")[["Author"]] `<br />
**Last modified:** `r file.info("MsBackendHmdb.Rmd")$mtime`<br />
**Compiled**: `r date()`

```{r, echo = FALSE, message = FALSE}
library(Spectra)
library(BiocStyle)
```

# Introduction

The `Spectra` package provides a central infrastructure for the handling of Mass
Spectrometry (MS) data. The package supports interchangeable use of different
*backends* to import MS data from a variety of sources (such as mzML files). The
`MsBackendHmdb` package enables, with the `MsBackendHmdbXml` object, import of
MS/MS spectrum data from xml files from The Human Metabolome Database
[HMDB](http://www.hmdb.ca). This vignette illustrates the usage of the
`MsBackendHmdb` package to enable HMDB data usage in `Spectra`.

# Importing MS/MS data from HMDB xml files

Spectral data from HMDB can be downloaded in xml format, one xml file per
spectrum. In our short example we load 4 such xml files which are provided with
this package. Below we first load all required libraries and define the file
names of the MS/MS spectra xml files.

```{r load-libs}
library(Spectra)
library(MsBackendHmdb)

fls <- dir(system.file("xml", package = "MsBackendHmdb"),
           full.names = TRUE, pattern = "xml$")
```

MS data can be accessed and analyzed through `Spectra` objects. Below we create
a `Spectra` with the data from the above xml files. To this end we provide the
file names and specify to use a `MsBackendHmdbXml()` backend as *source* to
enable data import.

```{r import}
sps <- Spectra(fls, source = MsBackendHmdbXml())
```

With that we have now full access to all imported spectra variables that we list
below.

```{r spectravars}
spectraVariables(sps)
```

Besides default spectra variables, such as `msLevel`, `rtime`, `precursorMz`
(most of which are however not defined in the xml files from HMDB) we have also
additional spectra variables such as the `spectrum_id` (the ID of the spectrum
in the HMDB database), `compound_id` (the metabolite identifier), `splash` or
`instrument_type`. Below we list the instrument type for the 4 spectra.

```{r instrument}
sps$instrument_type
```

The last spectrum was predicted and the instrument type is thus set to `NA`.

In addition we can also access the m/z and intensity values of each spectrum.

```{r mz}
mz(sps)
intensity(sps)
```

## Importing all MS/MS spectra from HMDB

It is also possible to import all MS/MS spectra from HMDB into a `Spectra`
object. For this we have to firstly download all *xml* files from the
*downloads* page of HMDB. HMDB allows to download all files in a single archive,
unzipping that will result in a folder with a very large number of small files,
one file per spectrum. Note also that this folder will contain also NMR and
other types of spectra. The variable `path` below is supposed to point to the
folder where all xml files can be found. We first list all xml files in that
folder using the pattern `"ms_ms_spectrum"` to get only file names for MS/MS
spectra.

```{r all-select-files, eval = FALSE}
path <- "~/data/hmdb_all_spectra/"
fls <- dir(path, pattern = "ms_ms_spectrum", full.names = TRUE)
```

With that we can now import the data and create a `Spectra` object representing
the collection of all HMDB MS2 spectra. Setting `nonStop = TRUE` prevents the
call to stop whenever it encounters problematic xml files (like xml files
without peaks). Note that the import of about 400,000 MS/MS spectra can take a
long time (in the range of one to several hours).

```{r all-import, eval = FALSE}
sps_hmdb <- Spectra(fls, source = MsBackendHmdbXml(), nonStop = TRUE)
```

# Session information

```{r}
sessionInfo()
```
